---
title: 'Predicting Income: EDA'
author: "Rick Fontenot"
date: "7/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(VIM)
library(caret)
library(corrplot)
library(ggplot2) 
library(ggthemes)
```

Load Theme for plots

```{r}
theme_set(theme_fivethirtyeight())
theme_update(axis.title = element_text()) #the default for fivethirtyeight is to not show axis labels, this removes that default so we can choose to specify and display axis titles
theme_update(plot.title = element_text(hjust = 0.5)) # changing default to center all titles
```

Load Data downloaded from UCI and stored on github repo
https://archive.ics.uci.edu/ml/datasets/Adult

```{r load data}
adult = read.csv("https://raw.githubusercontent.com/rickfontenot/Predicting_Income/main/adult.data", header = FALSE)
```

Description of variables from UCI:

Response: >50K, <=50K.

age: continuous.
workclass: Private, Self-emp-not-inc, Self-emp-inc, Federal-gov, Local-gov, State-gov, Without-pay, Never-worked.
fnlwgt: continuous.
education: Bachelors, Some-college, 11th, HS-grad, Prof-school, Assoc-acdm, Assoc-voc, 9th, 7th-8th, 12th, Masters, 1st-4th, 10th, Doctorate, 5th-6th, Preschool.
education-num: continuous.
marital-status: Married-civ-spouse, Divorced, Never-married, Separated, Widowed, Married-spouse-absent, Married-AF-spouse.
occupation: Tech-support, Craft-repair, Other-service, Sales, Exec-managerial, Prof-specialty, Handlers-cleaners, Machine-op-inspct, Adm-clerical, Farming-fishing, Transport-moving, Priv-house-serv, Protective-serv, Armed-Forces.
relationship: Wife, Own-child, Husband, Not-in-family, Other-relative, Unmarried.
race: White, Asian-Pac-Islander, Amer-Indian-Eskimo, Other, Black.
sex: Female, Male.
capital-gain: continuous.
capital-loss: continuous.
hours-per-week: continuous.
native-country: United-States, Cambodia, England, Puerto-Rico, Canada, Germany, Outlying-US(Guam-USVI-etc), India, Japan, Greece, South, China, Cuba, Iran, Honduras, Philippines, Italy, Poland, Jamaica, Vietnam, Mexico, Portugal, Ireland, France, Dominican-Republic, Laos, Ecuador, Taiwan, Haiti, Columbia, Hungary, Guatemala, Nicaragua, Scotland, Thailand, Yugoslavia, El-Salvador, Trinadad&Tobago, Peru, Hong, Holand-Netherlands.

Add column names to data set:
```{r}
# NOTE: names using underscore instead of hyphen so they can be referenced easier later
colnames(adult) <- c("age","workclass","fnlwgt","education","education_num","marital_status","occupation","relationship","race","sex","capital_gain","capital_loss","hours_per_week","native_country","income")

view(adult)
```

Investigate NA values to determine what needs resolution

```{r}
aggr_plot <- aggr(adult, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(adult), cex.axis=.7, gap=3, ylab=c("Percent data missing","Combinations Missing"))
#Note there are not missing values showing up initially but that's because the missing values are represented by "?" instead of NA

#Replace "?" with NA and re-do missing value analysis
adult[, 1:14][adult[, 1:14] == " ?"] <- NA
view(adult)

aggr_plot <- aggr(adult, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(adult), cex.axis=.7, gap=3, ylab=c("Percent data missing","Combinations Missing"))
marginplot(adult[c(2,7)])
marginplot(adult[c(2,14)])
marginplot(adult[c(7,14)])

#occupation missing 5.66% of values
#workclass missing 5.64% of values
#native-country missing 1.79& of values
#Note that half of the missing workclass values occur on observations that are also missing occupation
```

Examine formats of data available

```{r}
str(adult)

#Convert character vars to factors and make list of vars
adult$workclass <- as.factor(adult$workclass)
adult$education <- as.factor(adult$education)
adult$marital_status <- as.factor(adult$marital_status)
adult$occupation <- as.factor(adult$occupation)
adult$relationship <- as.factor(adult$relationship)
adult$race <- as.factor(adult$race)
adult$sex <- as.factor(adult$sex)
adult$native_country <- as.factor(adult$native_country)
adult$income <- as.factor(adult$income)

categorical.explanatory = c("workclass","education","marital_status","occupation","relationship","race","sex","native_country")

str(adult)

```


```{r}

```
