---
title: "Income_Simple_Model_2"
author: "Cameron Stewart"
date: "8/4/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE,message=FALSE}
library(glmnet)
library(ROCR)
library(MASS)
library(ggplot2)
library(pheatmap)
library(randomForest)
library(dplyr)
library(tidyverse)
library(VIM)
library(caret)
library(corrplot)
library(ggplot2) 
library(ggthemes)
library(vcd)
library(gtsummary)
library(Hmisc)
library(pROC)

```

Bring in train/test data:
```{r}
train_na_rm = read.csv("https://raw.githubusercontent.com/rickfontenot/Predicting_Income/main/data/Test_Train_Set/NAs_Removed/train_na_rm.csv", header = TRUE,stringsAsFactors = TRUE)

test_na_rm = read.csv("https://raw.githubusercontent.com/rickfontenot/Predicting_Income/main/data/Test_Train_Set/NAs_Removed/test_na_rm.csv", header = TRUE,stringsAsFactors = TRUE)

train_na_as_cat = read.csv("https://raw.githubusercontent.com/rickfontenot/Predicting_Income/main/data/Test_Train_Set/NAs_as_a_category/train_na_as_cat.csv", header = TRUE,stringsAsFactors = TRUE)

test_na_as_cat = read.csv("https://raw.githubusercontent.com/rickfontenot/Predicting_Income/main/data/Test_Train_Set/NAs_as_a_category/test_na_as_cat.csv", header = TRUE,stringsAsFactors = TRUE)
```

Add features from EDA to the train/test sets

```{r}

train_na_rm=train_na_rm[,-1]
test_na_rm=test_na_rm[,-1]
train_na_as_cat=train_na_as_cat[,-1]
test_na_as_cat=test_na_as_cat[,-1]

#train_na_rm
train_na_rm$income.binary<-0
train_na_rm$income.binary[train_na_rm$income==" >50K"] <- 1
train_na_rm$income.binary = as.factor(train_na_rm$income.binary)

train_na_rm$marriage_status <- "Other"
train_na_rm$marriage_status[train_na_rm$marital_status %in% c("Married-AF-spouse","Married-civ-spouse")] <- "Married"
train_na_rm$marriage_status[train_na_rm$marital_status %in% c("Divorced","Married-spouse-absent","Separated","Widowed")] <- "Previously-Married"
train_na_rm$marriage_status[train_na_rm$marital_status %in% c("Never-married")] <- "Single"

train_na_rm$collar <- "Other"
train_na_rm$collar[train_na_rm$occupation %in% c("Adm-clerical")] <- "White-support"
train_na_rm$collar[train_na_rm$occupation %in% c("Exec-managerial","Prof-specialty","Protective-serv","Sales","Tech-support
")] <- "White"
train_na_rm$collar[train_na_rm$occupation %in% c("Armed-Forces
","Craft-repair","Farming-fishing","Handlers-cleaners","Machine-op-inspct","Other-service","Priv-house-serv","Transport-moving")] <- "Blue"

train_na_rm$age2 = train_na_rm$age^2
train_na_rm$age3 = train_na_rm$age^3
train_na_rm$education_num2 = train_na_rm$education_num^2
train_na_rm$hours_per_week2 = train_na_rm$hours_per_week^2

train_na_rm$race <- as.factor(train_na_rm$race)
train_na_rm$sex <- as.factor(train_na_rm$sex)
train_na_rm$marriage_status <- as.factor(train_na_rm$marriage_status)
train_na_rm$collar <- as.factor(train_na_rm$collar)

#test_na_rm
test_na_rm$income.binary<-0
test_na_rm$income.binary[test_na_rm$income==" >50K"] <- 1
test_na_rm$income.binary = as.factor(test_na_rm$income.binary)

test_na_rm$marriage_status <- "Other"
test_na_rm$marriage_status[test_na_rm$marital_status %in% c("Married-AF-spouse","Married-civ-spouse")] <- "Married"
test_na_rm$marriage_status[test_na_rm$marital_status %in% c("Divorced","Married-spouse-absent","Separated","Widowed")] <- "Previously-Married"
test_na_rm$marriage_status[test_na_rm$marital_status %in% c("Never-married")] <- "Single"

test_na_rm$collar <- "Other"
test_na_rm$collar[test_na_rm$occupation %in% c("Adm-clerical")] <- "White-support"
test_na_rm$collar[test_na_rm$occupation %in% c("Exec-managerial","Prof-specialty","Protective-serv","Sales","Tech-support
")] <- "White"
test_na_rm$collar[test_na_rm$occupation %in% c("Armed-Forces
","Craft-repair","Farming-fishing","Handlers-cleaners","Machine-op-inspct","Other-service","Priv-house-serv","Transport-moving")] <- "Blue"

test_na_rm$age2 = test_na_rm$age^2
test_na_rm$age3 = test_na_rm$age^3
test_na_rm$education_num2 = test_na_rm$education_num^2
test_na_rm$hours_per_week2 = test_na_rm$hours_per_week^2

test_na_rm$race <- as.factor(test_na_rm$race)
test_na_rm$sex <- as.factor(test_na_rm$sex)
test_na_rm$marriage_status <- as.factor(test_na_rm$marriage_status)
test_na_rm$collar <- as.factor(test_na_rm$collar)

#train_na_as_cat
train_na_as_cat$income.binary<-0
train_na_as_cat$income.binary[train_na_as_cat$income==" >50K"] <- 1
train_na_as_cat$income.binary = as.factor(train_na_as_cat$income.binary)

train_na_as_cat$marriage_status <- "Other"
train_na_as_cat$marriage_status[train_na_as_cat$marital_status %in% c("Married-AF-spouse","Married-civ-spouse")] <- "Married"
train_na_as_cat$marriage_status[train_na_as_cat$marital_status %in% c("Divorced","Married-spouse-absent","Separated","Widowed")] <- "Previously-Married"
train_na_as_cat$marriage_status[train_na_as_cat$marital_status %in% c("Never-married")] <- "Single"

train_na_as_cat$collar <- "Other"
train_na_as_cat$collar[train_na_as_cat$occupation %in% c("Adm-clerical")] <- "White-support"
train_na_as_cat$collar[train_na_as_cat$occupation %in% c("Exec-managerial","Prof-specialty","Protective-serv","Sales","Tech-support
")] <- "White"
train_na_as_cat$collar[train_na_as_cat$occupation %in% c("Armed-Forces
","Craft-repair","Farming-fishing","Handlers-cleaners","Machine-op-inspct","Other-service","Priv-house-serv","Transport-moving")] <- "Blue"

train_na_as_cat$age2 = train_na_as_cat$age^2
train_na_as_cat$age3 = train_na_as_cat$age^3
train_na_as_cat$education_num2 = train_na_as_cat$education_num^2
train_na_as_cat$hours_per_week2 = train_na_as_cat$hours_per_week^2

train_na_as_cat$race <- as.factor(train_na_as_cat$race)
train_na_as_cat$sex <- as.factor(train_na_as_cat$sex)
train_na_as_cat$marriage_status <- as.factor(train_na_as_cat$marriage_status)
train_na_as_cat$collar <- as.factor(train_na_as_cat$collar)

#test_na_as_cat
test_na_as_cat$income.binary<-0
test_na_as_cat$income.binary[test_na_as_cat$income==" >50K"] <- 1
test_na_as_cat$income.binary = as.factor(test_na_as_cat$income.binary)

test_na_as_cat$marriage_status <- "Other"
test_na_as_cat$marriage_status[test_na_as_cat$marital_status %in% c("Married-AF-spouse","Married-civ-spouse")] <- "Married"
test_na_as_cat$marriage_status[test_na_as_cat$marital_status %in% c("Divorced","Married-spouse-absent","Separated","Widowed")] <- "Previously-Married"
test_na_as_cat$marriage_status[test_na_as_cat$marital_status %in% c("Never-married")] <- "Single"

test_na_as_cat$collar <- "Other"
test_na_as_cat$collar[test_na_as_cat$occupation %in% c("Adm-clerical")] <- "White-support"
test_na_as_cat$collar[test_na_as_cat$occupation %in% c("Exec-managerial","Prof-specialty","Protective-serv","Sales","Tech-support
")] <- "White"
test_na_as_cat$collar[test_na_as_cat$occupation %in% c("Armed-Forces
","Craft-repair","Farming-fishing","Handlers-cleaners","Machine-op-inspct","Other-service","Priv-house-serv","Transport-moving")] <- "Blue"

test_na_as_cat$age2 = test_na_as_cat$age^2
test_na_as_cat$age3 = test_na_as_cat$age^3
test_na_as_cat$education_num2 = test_na_as_cat$education_num^2
test_na_as_cat$hours_per_week2 = test_na_as_cat$hours_per_week^2

test_na_as_cat$race <- as.factor(test_na_as_cat$race)
test_na_as_cat$sex <- as.factor(test_na_as_cat$sex)
test_na_as_cat$marriage_status <- as.factor(test_na_as_cat$marriage_status)
test_na_as_cat$collar <- as.factor(test_na_as_cat$collar)
```

Adding my custom predictors
```{r}

for (i in 1:4){
  if (i==1){  
    #Transform holder allows me to copy and paste transform code to multiple test and training sets with ease
    transform_holder=train_na_rm
  }
  if (i==2){  
    #Transform holder allows me to copy and paste transform code to multiple test and training sets with ease
    transform_holder=test_na_rm
  }
  if (i==3){  
    #Transform holder allows me to copy and paste transform code to multiple test and training sets with ease
    transform_holder=train_na_as_cat
  }
  if (i==4){  
    #Transform holder allows me to copy and paste transform code to multiple test and training sets with ease
    transform_holder=test_na_as_cat
  }
  
transform_holder$work_sector <- "Other"
transform_holder$work_sector[transform_holder$workclass %in% c("Federal-gov","Local-gov","State-gov")] <- "Government"
transform_holder$work_sector[transform_holder$workclass %in% c("Private")] <- "Private"
transform_holder$work_sector[transform_holder$workclass %in% c("Self-emp-inc","Self-emp-not-inc")] <- "Self_Employed"
transform_holder$work_sector[transform_holder$workclass %in% c("Never-worked","Without-pay")] <- "Not_Working"
transform_holder$work_sector = as.factor(transform_holder$work_sector)
  
  #### My Additional Releveling
  
  #Capital Gains
  transform_holder$cap_gains_bin=ifelse(transform_holder$capital_gain>0,'Gain','No Gain')
  transform_holder$cap_gains_bin=factor(transform_holder$cap_gains_bin)
  
  #Capital Losses
  transform_holder$cap_loss_bin=ifelse(transform_holder$capital_loss>0,'Loss','No Loss')
  transform_holder$cap_loss_bin=factor(transform_holder$cap_loss_bin)
  
  ##### My Experimental Re-leveling
  
  ## The purpose is to maximize predictability
  ##My plan is to test for significance differences for categories with less than 5% of total population. Then combine non-significant differences
  
  #Occupation
  
  transform_holder$alt_occupation <- as.character(transform_holder$occupation)
  transform_holder$alt_occupation[transform_holder$occupation %in% c("Tech-support","Protective-serv","Armed-Forces")] <- "Tech_and_Protection"
  transform_holder$alt_occupation[transform_holder$occupation %in% c("Farming-fishing","Machine-op-inspct")] <- "Agr_and_Mach_Ops"
  transform_holder$alt_occupation[transform_holder$occupation %in% c("Other-service","Priv-house-serv")] <- "Other-service"
  transform_holder$alt_occupation = factor(transform_holder$alt_occupation)
  
    #Combine Tech-Support, Protective-serv, and Armed forces
    #Combine Farming-fishing and Machine-op-inspect
    #Combine Other-Service and Priv-house-serv
  
  #sort(summary(transform_holder$alt_occupation)/dim(transform_holder)[1],decreasing = TRUE)
  
  #occ_model<-glm(income~relevel(alt_occupation, ref = "Tech_and_Protection"),family="binomial",data=transform_holder)
  #summary(occ_model)
  
  #Marital Status
  
  transform_holder$alt_marital_status <- as.character(transform_holder$marital_status)
  transform_holder$alt_marital_status[transform_holder$marital_status %in% c("Separated","Married-spouse-absent")] <- "sep_and_abs_spouse"
  transform_holder$alt_marital_status[transform_holder$marital_status %in% c("Widowed","Divorced")] <- "Widowed_Divorced"
  transform_holder$alt_marital_status[transform_holder$marital_status %in% c("Married-AF-spouse","Married-civ-spouse")] <- "Married_with_spouse"
  transform_holder$alt_marital_status = factor(transform_holder$alt_marital_status)
  
    #Combine Separated and Married-spouse-absent
    #Combine Widowed and Divorced
    #Combine Married-AF-spouse and Married-civ-spouse
  
  #sort(summary(transform_holder$alt_marital_status)/dim(transform_holder)[1],decreasing = TRUE)
  
  #ms_model<-glm(income~relevel(alt_marital_status, ref = "Married_with_spouse"),family="binomial",data=transform_holder)
  #summary(ms_model)
  
  
  #Workclass
  
  transform_holder$alt_workclass <- as.character(transform_holder$workclass)
  transform_holder$alt_workclass[transform_holder$workclass %in% c("Without-pay","Never-worked","no_response")] <- "Other"
  transform_holder$alt_workclass[transform_holder$workclass %in% c("State-gov","Self-emp-not-inc")] <- "state_gov_and_self_emp_non_inc"
  transform_holder$alt_workclass = factor(transform_holder$alt_workclass)
  
    #Combine Without-pay and Never-worked and no_response as other
    #Combine State-gov and Self-emp-not-inc
  
  #sort(summary(transform_holder$alt_workclass)/dim(transform_holder)[1],decreasing = TRUE)
  
  #wc_model<-glm(income~relevel(alt_workclass, ref = "Federal-gov"),family="binomial",data=transform_holder)
  #summary(wc_model)
  
  #Relationship
  
  #transform_holder$alt_relationship <- as.character(transform_holder$relationship)
  #transform_holder$alt_relationship = factor(transform_holder$alt_relationship)
  
    #Left this one the same because all were significantly different
  
  #sort(summary(transform_holder$alt_relationship)/dim(transform_holder)[1],decreasing = TRUE)
  
  #wc_model<-glm(income~relevel(alt_relationship, ref = "Wife"),family="binomial",data=transform_holder)
  #summary(wc_model)
  
  #Race
  
  transform_holder$alt_race <- as.character(transform_holder$race)
  transform_holder$alt_race[transform_holder$race %in% c("White","Asian-Pac-Islander")] <- "White_and_API"
  transform_holder$alt_race[transform_holder$race %in% c("Amer-Indian-Eskimo","Other")] <- "Other"
  transform_holder$alt_race = factor(transform_holder$alt_race)
  
    #Combine White with Asian-Pac-Islander
    #Combine Amer-Indian-Eskimo and Other
    #Made decision not to combine Black and Other despite in-significant difference
  
  #sort(summary(transform_holder$alt_race)/dim(transform_holder)[1],decreasing = TRUE)
  
  #race_model<-glm(income~relevel(alt_race, ref = "Other"),family="binomial",data=transform_holder)
  #summary(race_model)
  
  #Native Country
  
  transform_holder$alt_native_country <- as.character(transform_holder$native_country)
  transform_holder$alt_native_country[!(transform_holder$native_country %in% c("United-States","Mexico","Philippines","no_response"))] <- "Other"
  transform_holder$alt_native_country[transform_holder$native_country %in% c("United-States","no_response")] <- "US_and_nr"
  transform_holder$alt_native_country = factor(transform_holder$alt_native_country)
  
    #Combine all non US, Mex, and Philippines countries into other
    #Combine US and no_response
    #Made decision not to combine Black and Other despite in-significant difference
  
  #sort(summary(transform_holder$alt_native_country)/dim(transform_holder)[1],decreasing = TRUE)
  
  #nc_model<-glm(income~relevel(alt_native_country, ref = "Philippines"),family="binomial",data=transform_holder)
  #summary(nc_model)
  
  
  #Education
  transform_holder$alt_education <- as.character(transform_holder$education)
  transform_holder$alt_education[transform_holder$education %in% c("Assoc-acdm","Assoc-voc","Preschool")] <- "Assoc_and_Prek"
  transform_holder$alt_education[transform_holder$education %in% c("9th","7th-8th","11th","5th-6th","1st-4th","10th")] <- "1st-11th"
  transform_holder$alt_education[transform_holder$education %in% c("Doctorate","Prof-school")] <- "Doctorate and Prof School"
  transform_holder$alt_education = factor(transform_holder$alt_education)
  
    #Combine Assoc-acdm and Assoc-voc and Preschool
    #Combine 1st - 11th
    #Combine Doctorate and Prof-school
  
  #sort(summary(transform_holder$alt_education)/dim(transform_holder)[1],decreasing = TRUE)
  
  #edu_model<-glm(income~relevel(alt_education, ref = "12th"),family="binomial",data=transform_holder)
  #summary(edu_model)

  if (i==1){  
    #Transform holder allows me to copy and paste transform code to multiple test and training sets with ease
    train_na_rm=transform_holder
  }
  if (i==2){  
    #Transform holder allows me to copy and paste transform code to multiple test and training sets with ease
    test_na_rm=transform_holder
  }
  if (i==3){  
    #Transform holder allows me to copy and paste transform code to multiple test and training sets with ease
    train_na_as_cat=transform_holder
  }
  if (i==4){  
    #Transform holder allows me to copy and paste transform code to multiple test and training sets with ease
    test_na_as_cat=transform_holder
  }
  
}
```

Create Feature lists for models

```{r}
features.simple = c("income","age","education_num","race","sex","capital_gain","capital_loss","hours_per_week","marriage_status","collar")

features.complex = c("income","age","education_num","race","sex","capital_gain","capital_loss","hours_per_week","marriage_status","collar","age2","age3","education_num2","hours_per_week2")

original.variables <- c("income","age","workclass","fnlwgt","education","education_num","marital_status","occupation","relationship","race","sex","capital_gain","capital_loss","hours_per_week","native_country")
```

Base Logistic regression model with all original variables for comparison

```{r}
dat.train.x <- train_na_rm %>% select(all_of(original.variables))
dat.train.x <- model.matrix(income~.,dat.train.x)
dat.train.y <- train_na_rm$income

#glmnet requires a matrix 
cvfit <- cv.glmnet(dat.train.x, dat.train.y, family = "binomial", type.measure = "class", nlambda = 100)
plot(cvfit)
coef(cvfit, s = "lambda.min")

#Get Test Set
dat.val1.x <- test_na_rm %>% select(all_of(original.variables))
dat.val1.x <- model.matrix(income~.,dat.val1.x)
dat.val1.y <- test_na_rm$income


#Run model from training set on valid set I
fit.pred1 <- predict(cvfit, newx = dat.val1.x, type = "response")

#ROC curves
pred1 <- prediction(fit.pred1[,1], dat.val1.y)
roc.perf1 = performance(pred1, measure = "tpr", x.measure = "fpr")
auc.val1 <- performance(pred1, measure = "auc")
auc.val1 <- auc.val1@y.values
plot(roc.perf1)
abline(a=0, b= 1)
text(x = .40, y = .6,paste("AUC = ", round(auc.val1[[1]],3), sep = ""))
```

Run Simple Logistic regression with NA removed, add AUC metric

```{r}
dat.train.x <- train_na_rm %>% select(all_of(features.simple))
dat.train.x <- model.matrix(income~.,dat.train.x)
dat.train.y <- train_na_rm$income

#glmnet requires a matrix 
cvfit <- cv.glmnet(dat.train.x, dat.train.y, family = "binomial", type.measure = "class", nlambda = 100)
plot(cvfit)
coef(cvfit, s = "lambda.min")

#Get Test Set
dat.val1.x <- test_na_rm %>% select(all_of(features.simple))
dat.val1.x <- model.matrix(income~.,dat.val1.x)
dat.val1.y <- test_na_rm$income


#Run model from training set on valid set I
fit.pred1 <- predict(cvfit, newx = dat.val1.x, type = "response")

#ROC curves
pred1 <- prediction(fit.pred1[,1], dat.val1.y)
roc.perf1 = performance(pred1, measure = "tpr", x.measure = "fpr")
auc.val1 <- performance(pred1, measure = "auc")
auc.val1 <- auc.val1@y.values
plot(roc.perf1)
abline(a=0, b= 1)
text(x = .40, y = .6,paste("AUC = ", round(auc.val1[[1]],3), sep = ""))
```

Run Simple Logistic regression with NA as category, add AUC metric

```{r}
dat.train.x <- train_na_as_cat %>% select(all_of(features.simple))
dat.train.x <- model.matrix(income~.,dat.train.x)
dat.train.y <- train_na_as_cat$income

#glmnet requires a matrix 
cvfit <- cv.glmnet(dat.train.x, dat.train.y, family = "binomial", type.measure = "class", nlambda = 100)
plot(cvfit)
coef(cvfit, s = "lambda.min")

#Get Test Set
dat.val1.x <- train_na_as_cat %>% select(all_of(features.simple))
dat.val1.x <- model.matrix(income~.,dat.val1.x)
dat.val1.y <- train_na_as_cat$income


#Run model from training set on valid set I
fit.pred1 <- predict(cvfit, newx = dat.val1.x, type = "response")

#ROC curves
pred1 <- prediction(fit.pred1[,1], dat.val1.y)
roc.perf1 = performance(pred1, measure = "tpr", x.measure = "fpr")
auc.val1 <- performance(pred1, measure = "auc")
auc.val1 <- auc.val1@y.values
plot(roc.perf1)
abline(a=0, b= 1)
text(x = .40, y = .6,paste("AUC = ", round(auc.val1[[1]],3), sep = ""))
```

Testing Rick's features.simple on both data sets

```{r}
variable.names(train_na_as_cat)

features.simple = c("income","age","education_num","race","sex","capital_gain","capital_loss","hours_per_week","marriage_status","collar")

simp1=train_na_as_cat %>% select(all_of(features.simple))
#simp2=train_na_rm %>% select(all_of(features.simple))

#fit.na_as_cat<-glm(income~.,family="binomial",data=simp1)
DF <- within(simp1, race <- relevel(race, ref = "White"))
fit.na_as_cat<-glm(income~.,family="binomial",data=DF)

levels(train_na_as_cat$race)

predprobs<-predict(fit.na_as_cat,test_na_as_cat,type="response")
log.roc<-roc(response=test_na_as_cat$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

#Manually adjusting threshhold
threshold=.2176
log.predclass<-ifelse(predprobs>threshold,">50K","<=50K")
log.predclass<-factor(log.predclass)
confusionMatrix(log.predclass,test_na_as_cat$income)

exp(cbind("Odds ratio" = coef(fit.na_as_cat), confint.default(fit.na_as_cat, level = 0.95)))

summary(train_na_as_cat)

# fit.na_rm<-glm(income~.,family="binomial",data=simp2)
# predprobs<-predict(fit.na_rm,test_na_rm,type="response")
# log.roc<-roc(response=test_na_rm$income,predictor=predprobs,levels=c("<=50K",">50K"))
# plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
# auc(log.roc)

summary(fit.na_as_cat)
#car::residualPlots(fit.na_as_cat)
car::influenceIndexPlot(fit.na_as_cat,id.n=3)
car::influencePlot(fit.na_as_cat,col="red",id.n=3)
fit.na_as_cat_no_outliers=update(fit.na_as_cat,subset=c(-20872))
car::compareCoefs(fit.na_as_cat_no_outliers,fit.na_as_cat,se = FALSE)

features.numerical = c("income","age","education_num","capital_gain","capital_loss","hours_per_week")

vif=car::vif(fit.na_as_cat)
vif[,3]^2


```


Looking into cross validation

```{r}
# # define training control
# train_control <- trainControl(method = "cv", number = 10)
# 
# # train the model on training set
# model <- train(income ~ .,
#                data = DF,
#                trControl = train_control,
#                method = "glm",
#                family=binomial())


```

Testing more complex set

```{r}
variable.names(train_na_as_cat)

features.complex = c("income","age","education_num","race","sex","capital_gain","capital_loss","hours_per_week","marriage_status","collar","hours_per_week2","age2","age3","education_num2")

simp1=train_na_as_cat %>% select(all_of(features.complex))

DF <- within(simp1, race <- relevel(race, ref = "White"))
fit.na_as_cat<-glm(income~.,family="binomial",data=DF)

predprobs<-predict(fit.na_as_cat,test_na_as_cat,type="response")
log.roc<-roc(response=test_na_as_cat$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

#Manually adjusting threshhold
threshold=.2176
log.predclass<-ifelse(predprobs>threshold,">50K","<=50K")
log.predclass<-factor(log.predclass)
confusionMatrix(log.predclass,test_na_as_cat$income)

exp(cbind("Odds ratio" = coef(fit.na_as_cat), confint.default(fit.na_as_cat, level = 0.95)))

summary(fit.na_as_cat)


```

Adding Native Country

```{r}
features.simple = c("income","age","education_num","race","sex","capital_gain","capital_loss","hours_per_week","marriage_status","collar","alt_native_country")

simp1=train_na_as_cat %>% select(all_of(features.simple))
simp2=train_na_rm %>% select(all_of(features.simple))

fit.na_as_cat<-glm(income~.,family="binomial",data=simp1)
predprobs<-predict(fit.na_as_cat,test_na_as_cat,type="response")
log.roc<-roc(response=test_na_as_cat$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_as_cat)

fit.na_rm<-glm(income~.,family="binomial",data=simp2)
predprobs<-predict(fit.na_rm,test_na_rm,type="response")
log.roc<-roc(response=test_na_rm$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_rm)
```

Adding alt_Native_Country
```{r}
features.simple = c("income","age","education_num","race","sex","capital_gain","capital_loss","hours_per_week","marriage_status","collar","alt_native_country")

simp1=train_na_as_cat %>% select(all_of(features.simple))
simp2=train_na_rm %>% select(all_of(features.simple))

fit.na_as_cat<-glm(income~.,family="binomial",data=simp1)
predprobs<-predict(fit.na_as_cat,test_na_as_cat,type="response")
log.roc<-roc(response=test_na_as_cat$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)


fit.na_rm<-glm(income~.,family="binomial",data=simp2)
predprobs<-predict(fit.na_rm,test_na_rm,type="response")
log.roc<-roc(response=test_na_rm$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_as_cat)
summary(fit.na_rm)

#fit.na_rm resulted in the same AUC of .8976 and the same AIC of 16239 so we will not add alt_native_country
```

Adding alt_marital_status

```{r}
variable.names(train_na_as_cat)

features.simple = c("income","age","education_num","race","sex","capital_gain","capital_loss","hours_per_week","alt_marital_status","collar")

simp1=train_na_as_cat %>% select(all_of(features.simple))
simp2=train_na_rm %>% select(all_of(features.simple))

fit.na_as_cat<-glm(income~.,family="binomial",data=simp1)
predprobs<-predict(fit.na_as_cat,test_na_as_cat,type="response")
log.roc<-roc(response=test_na_as_cat$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_as_cat)

fit.na_rm<-glm(income~.,family="binomial",data=simp2)
predprobs<-predict(fit.na_rm,test_na_rm,type="response")
log.roc<-roc(response=test_na_rm$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_rm)

#Same AUC and worse AIC so will not use
```

Adding alt_occupation

```{r}
variable.names(train_na_as_cat)

features.simple = c("income","age","education_num","race","sex","capital_gain","capital_loss","hours_per_week","marriage_status","alt_occupation")

simp1=train_na_as_cat %>% select(all_of(features.simple))
simp2=train_na_rm %>% select(all_of(features.simple))

fit.na_as_cat<-glm(income~.,family="binomial",data=simp1)
predprobs<-predict(fit.na_as_cat,test_na_as_cat,type="response")
log.roc<-roc(response=test_na_as_cat$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_as_cat)

fit.na_rm<-glm(income~.,family="binomial",data=simp2)
predprobs<-predict(fit.na_rm,test_na_rm,type="response")
log.roc<-roc(response=test_na_rm$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_rm)

#Better AUC and AIC so keeping alt occupation
```

Adding captial bins

```{r}
variable.names(train_na_as_cat)

features.simple = c("income","age","education_num","race","sex","cap_gains_bin","cap_loss_bin","hours_per_week","marriage_status","alt_occupation")

simp1=train_na_as_cat %>% select(all_of(features.simple))
simp2=train_na_rm %>% select(all_of(features.simple))

fit.na_as_cat<-glm(income~.,family="binomial",data=simp1)
predprobs<-predict(fit.na_as_cat,test_na_as_cat,type="response")
log.roc<-roc(response=test_na_as_cat$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_as_cat)

fit.na_rm<-glm(income~.,family="binomial",data=simp2)
predprobs<-predict(fit.na_rm,test_na_rm,type="response")
log.roc<-roc(response=test_na_rm$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_rm)

#Worse AUC and AIC so not keeping capital gain and loss bins only
```

Adding capital bins and capital results

```{r}
variable.names(train_na_as_cat)

features.simple = c("income","age","education_num","race","sex","capital_gain","capital_loss","cap_gains_bin","cap_loss_bin","hours_per_week","marriage_status","alt_occupation")

simp1=train_na_as_cat %>% select(all_of(features.simple))
simp2=train_na_rm %>% select(all_of(features.simple))

fit.na_as_cat<-glm(income~.,family="binomial",data=simp1)
predprobs<-predict(fit.na_as_cat,test_na_as_cat,type="response")
log.roc<-roc(response=test_na_as_cat$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_as_cat)

fit.na_rm<-glm(income~.,family="binomial",data=simp2)
predprobs<-predict(fit.na_rm,test_na_rm,type="response")
log.roc<-roc(response=test_na_rm$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_rm)

#Better AUC and AIC so keeping bins and totals in the model
```

Trying alt race

```{r}
variable.names(train_na_as_cat)

features.simple = c("income","age","education_num","alt_race","sex","capital_gain","capital_loss","cap_gains_bin","cap_loss_bin","hours_per_week","marriage_status","alt_occupation")

simp1=train_na_as_cat %>% select(all_of(features.simple))
simp2=train_na_rm %>% select(all_of(features.simple))

fit.na_as_cat<-glm(income~.,family="binomial",data=simp1)
predprobs<-predict(fit.na_as_cat,test_na_as_cat,type="response")
log.roc<-roc(response=test_na_as_cat$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_as_cat)

fit.na_rm<-glm(income~.,family="binomial",data=simp2)
predprobs<-predict(fit.na_rm,test_na_rm,type="response")
log.roc<-roc(response=test_na_rm$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_rm)

#No change in AUC or AIC so not keeping alt_race
```


Trying alt race

```{r}
variable.names(train_na_as_cat)

features.simple = c("income","age","education_num","race","sex","capital_gain","capital_loss","cap_gains_bin","cap_loss_bin","hours_per_week","marriage_status","alt_occupation","alt_workclass")

simp1=train_na_as_cat %>% select(all_of(features.simple))
simp2=train_na_rm %>% select(all_of(features.simple))

fit.na_as_cat<-glm(income~.,family="binomial",data=simp1)
predprobs<-predict(fit.na_as_cat,test_na_as_cat,type="response")
log.roc<-roc(response=test_na_as_cat$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_as_cat)

fit.na_rm<-glm(income~.,family="binomial",data=simp2)
predprobs<-predict(fit.na_rm,test_na_rm,type="response")
log.roc<-roc(response=test_na_rm$income,predictor=predprobs,levels=c("<=50K",">50K"))
plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
auc(log.roc)

summary(fit.na_rm)
#Improved AUC and AIC so keeping alt_workclass
```